On the characteristic and diameter of planar integral point sets \footnote{ The work of the first author was partially supported by the Theoretical Physics and Mathematics Advancement Foundation “BASIS”. }

### 2025-02-22 char-4k-bound/char-4k-bound.tex
[4k3] Add a note about the asterisk notation
[4k3] Remove a misleading comma in 10 655
[4k3] Fix two remarks from the Referee

\begin{definition}
	\label{def:char}
	The characteristic of a planar IPS $M$ is a squarefree number $q$
	such that the area of any triangle $M_1M_2M_3$, $\{M_1,M_2,M_3\}\subset M$,
	is commensurable with $\sqrt{q}$.
	Notation: $\chr M = q$.
\end{definition}

@@

\begin{definition}
	Let $M_1 M_2$ be an edge of $M\in \mathfrak{M}$.
	For $N\in\mathbb N_\pm$, $|n| < |M_1 M_2|$
	we will say that the set of points
	\begin{equation}
		\{M_0 : |M_0 M_1| - |M_0 M_2| = n\}
	\end{equation}
	is called the \emph{$n$-th Erdős curve}.
\end{definition}

@@

For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018-translit}:
$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
which means that each abscissa is multiplied by $1/q$
and each ordinate is multiplied by $\sqrt{p}/q$,  i.e.
\begin{equation}
	\label{eq:char_lattice}
	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
	=
	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}
	,
\end{equation}
where $q$ is the characteristic of the IPS.
Such notation is made possible by the Grid Theorem.

@@

	According to the results~\eqref{d}, for $3 \leq n \leq 36$ the theorem is true.
	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20\ 000$  and our estimate on the diameter is also true.
	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10,655$.
	This estimate is weaker than the available numerical results. Therefore, from now on, we can assume that $n \geq 7^4$.

>>>>

\begin{definition}
	\label{def:char}
	The characteristic of a planar IPS $M$ is a squarefree number $q$
	such that the area of any triangle $M_1M_2M_3$, $\{M_1,M_2,M_3\}\subset M$,
	is commensurable with $\sqrt{q}$.
	Notation: $\chr M = q$.
\end{definition}
The characteristic of an IPS always exists and is unique.

@@

\begin{definition}
	Let $M_1 M_2$ be an edge of $M\in \mathfrak{M}$.
	For an integer $n$, $|n| < |M_1 M_2|$,
	we will call the set of points
	\begin{equation}
		\{M_0 : |M_0 M_1| - |M_0 M_2| = n\}
	\end{equation}
	the \emph{$n$-th Erdős curve}.
\end{definition}

@@

For convenience, we use the notation \cite{our-ped-2018,our-pmm-2018,our-vmmsh-2018-translit}:
$\sqrt{p}/q * \{ (x_1,y_1), ...,$ $ (x_n, y_n)  \}$,
which means that each abscissa is multiplied by $1/q$
and each ordinate is multiplied by $\sqrt{p}/q$,  i.e.
\begin{equation}
	\label{eq:char_lattice}
	\sqrt{p}/q * \{ (x_1,y_1), ..., (x_n, y_n)  \}
	=
	\left\{ \left(\frac{x_1}{q},\frac{y_1\sqrt{p}}{q}\right), ..., \left(\frac{x_n}{q},   \frac{y_n\sqrt{p}}{q}\right)  \right\}
	,
\end{equation}
where $q$ is the characteristic of the IPS.
Such notation is made possible by the Grid Theorem.
Note that we use the asterisk instead of the centered dot to emphasize
that we do not mean the usual coordinatewise multiplication;
instead, the denominator is appended to both coordinates,
but the square root multiplier is appended only to the second one.

@@

	According to the results~\eqref{d}, for $3 \leq n \leq 36$ the theorem is true.
	Moreover, it is known that for all $37 \leq n \leq 7^4$, we have $\overline{d}(2, n) > 20\ 000$  and our estimate on the diameter is also true.
	Indeed, $\overline{d}(2, 7^4) \geq \left(\frac{25}{36}\cdot 7^4\right)^{5/4} \approx 10 655$.
	This estimate is weaker than the available numerical results. Therefore, from now on, we can assume that $n \geq 7^4$.
